### 2017-09-14 papers/BijectiveLensSynth/popl_18_rebuttal.txt

The algorithm described includes all but compositional synthesis, which uses
predefined lenses.  To use a predefined lens, l : A <-> B, the implementation
develops equivalence classes between user defined data types with lenses between
them (in the above example, A and B would be placed in the same equivalence
class).  Next, a representative element of each equivalence class is chosen.
All user defined data types are then transformed into their representative, and
the string examples are converted into examples in the language of the
representative.

                                -------------

     - p21: Fig 9 "Number of expansions" --> "Number of solutions"??

Number of expansions is the correct phrase.  Figure 9 shows how much
nondeterministic search through expansions is required before RigidSynth
terminates.  In particular, this shows that full expansion inference performs
significantly better than the partial one that only performs the expansions it
knows it must perform.

@@

                                -------------

     - p21.  Also interesting to evaluate would be how time / number of
       solutions varies (e.g. for the full algorithm) as the number of
       examples provided varies.  Is the number of examples small?

The number of examples provided doesn't typically have an impact on runtime
speed, only correctness.  The number of valid solutions does vary with the
number of examples provided.  Typically, a small number of examples can greatly
reduce the number of possible solutions.  The number of examples is indeed
small, as is shown in Figure 11.

@@

      let bibtex_to_endnote_au =
        synth
          BIBTEXAUTHORINFO <=> TAGGEDAUTHORDEFNS

      let bibtex_to_endnote_au_synthesized_result
        : (lens in BIBTEXAUTHORINFO <=> TAGGEDAUTHORDEFNS) =
        del "author={"
          . ins "au - "
          . lens_swap (NAME . del ",") (lens_swap WSP NAME)* 
          . (del " and "
               . ins "\n au - "
               . lens_swap (NAME . del ",") (lens_swap WSP NAME)* )* 
          . del "}"

      let bibtex_to_endnote_au_human_written
        : (lens in BIBTEXAUTHORINFO <=> TAGGEDAUTHORDEFNS) =
        let single_author_convert =
          ins " au - "
            . lens_swap
                (NAME . del ",")
                (lens_swap WSP NAME)*
        in
        del "author={"
          . single_author_convert
          . (del " and "
               . ins "\n"
               . single_author_convert)* 
          . del "}"

Even the relatively simple lens that permutes the 3 elements needed for
single_author_convert is fairly complex.  The complexity becomes even more
apparent when working with large lenses comprised of many sub-lenses (though for
brevity we give here the small example above).  Lenses provide great power with
their invertibility guarantees, but they come at the cost of thinking about many
fiddly details when writing the terms.  Worrying about these fiddly details
while ALSO thinking about unambiguity restrictions is a very difficult task!

@@

If someone provides Sigma* and Sigma* as domains and ranges to optician it
will merely return the identity between those two.  If a user presents Sigma*
and Sigma* with some examples, we expect it will unroll the iterations until
the number of characters in the largest example has been enumerated, and
return something close to identity, but performing differently on those
examples, and some others to handle the readjustment.

                                -------------

     - Also, a better explanation of the difference between the DSLs of Lenses
       and Flash fill is required. For example, FlashFill can reverse a string,
       but the proposed language can’t.

The Boomerang lenses that we synthesize don't currently provide such
capabilities.  However, there is no technical reason this can't be done by
adding in a reverse-and-iterate operator to Boomerang.  Only slight adjustments
to the synthesis algorithm would be needed.

@@

On average, the domain and range specs are slightly larger than the lens itself.
See above for the chart and a commentary explaining why we believe writing
lenses is harder than writing regular expressions, despite the fact that lenses
alone are sometimes shorter than regular expression specifications.

                                -------------

     - What is the difference in size between manually written lenses and
       synthesized ones?

We have not manually written lenses for all our benchmarks.  For the ones we
have, we find that the sizes are fairly comparable, but the hand-written lenses
are more modular: with common portions factored out and given good names.

@@

                                -------------

     - Clarify how the benchmarks were selected and whether any were not
       selected a priori. Especially the flash* ones. Did you omit those that
       used character transformations such as upper-case?

We used benchmarks from three sources: Augeas benchmarks, FlashFill benchmarks,
and custom benchmarks.  We used Augeas benchmarks because it is a lens-based
tool that has seen industrial use.  We used FlashFill benchmarks because it is
a well-known and popular string transformation synthesis system.
We used custom benchmarks to find good uses for the tool in areas not previously
explored (like converting user address information from a CSV format into a
letter format).

For our Augeas benchmarks, we began synthesizing lenses for arbitrary problems
in the directory.  After a few of these, to show we weren't cherry-picking
lenses that our system could handle, we began turning the lenses into synthesis
problems alphabetically - first the As, then the Bs, etc.

@@

const("<Field Id=2></Field>","")
  | (const("<Field Id=2>","Title: ")
      . Id(text_char)
      . const("","")
      . iterate(const("","") . Id(text_char) . const("",""))
      . const("","")
      . const("</Field>",", "))

With minimization, the title field transformation would appear as:

const("<Field Id=2>","")
  . (Id("")
       | (Id(text_char) . iterate(Id(text_char)) . const("", ",")))
  . const("</Field>",""))

In particular, the useless concatenations of const("","") are removed and the
field boilerplate is shared rather than being duplicated in the two cases of the
disjunct.  Furthermore, when const(s,s) is present, it is converted into Id(s).

@@

We support the bijective subset given by the combinators in Section 2.  See the
general responses for a discussion on the usefulness of these combinators.

                                -------------

     - I was also a bit confused about the preliminaries in Sec. 2. In
       particular, it is not clear whether all or most of those preliminaries
       are from related work (but there are no references) or whether they are
       contributions of this paper.

The preliminaries define a specific subset of Boomerang that we aim to
synthesize.  Boomerang supports a variety of lenses - asymmetric lenses,
quotient lenses, and alignment lenses defined over a series of papers
[Combinators for bi-directional tree transformations: a linguistic approach to
the view update problem, POPL 2005; Boomerang: Resourceful Lenses for String
Data, POPL 2008; Quotient Lenses, ICFP 2008]. While this specific sublanguage
does not appear verbatim in a prior paper, it is a natural core language for
string lenses.

### 2017-09-14 papers/BijectiveLensSynth/popl_18_rebuttal.txt

Thanks for your comments!

We begin with brief responses to the most significant concerns, followed by a
(long and optional) appendix addressing all questions in detail.

@@

There are lens theories developed to handle synchronization tasks for data
formats not kept in bijective correspondence (quotient lenses to handle Category
1 non-bijectivity, and symmetric lenses to handle Category 2 non-bijectivity).
These theories have a bijective core, surrounded by either canonizers (for
quotient lenses) or projections (for symmetric lenses).  We believe that our
system can be used as a drop-in component for synthesis algorithms on these more
complex lens structures.

Reviewer 1 asked us directly about the potential for future work.
Along those lines, we have already implemented a system that
synthesizes quotient lenses from regular expressions augmented with equivalence
relations that solve the whitespace bijection problem.  This system
uses the algorithm provided in this paper as a drop-in component without change.
We conjecture the system could be augmented in a similar way with projections as
well.  We are not suggesting adding more results of this form to the current paper
because it is already very dense.

     - After having written the regular expressions, is it very difficult
       for users to then write the lens?  Are we saving anybody any
       significant effort?

While we still require users to write regular expressions, programming
regular expressions is easier than programming lenses:

@@

* Regular expressions are widely understood.  Few programmers know about
  programming with lenses.

* A regular expression describes one data format, while lenses deal with two
  at the same time.

* In the (common) case where the two formats have Kleene stars in different
  places, manual lens programming requires mentally transforming both to a
  common "aligned" form.  Our synthesis algorithm does this alignment
  automatically.  This alignment can be quite challenging for programmers.

@@

Engineering best practices dictate that we annotate the lenses with their
types.  While Boomerang does not need these annotations - being able to
infer the types from the terms - the types of the term serve as
documentation for future programmers to understand what formats the lens
maps between.  Furthermore, it provides an additional resilience to future
modifications of the lens, as it validates that the lens maps exactly
between strings of the provided formats - invalid inputs are not accepted
nor are the outputs ever invalid.  In this way, we are not requiring
additional work by the programmer to use our system; we are allowing the
programmer to save the difficulties of writing the term by using work they
should already be performing.

### 2017-09-14 papers/BijectiveLensSynth/popl_18_rebuttal.txt

Formally, they are incomparable: each can synthesize some transformations
that the other cannot.

Since we restrict ourselves to bijections, we are unable to express functions
that don't preserve data, such as
  - mapping from a format with whitespace to one without whitespace,
  - turning a mix of uppercase and lowercase characters into purely
    uppercase characters, and
  - extracting a file name from a file path.
FlashFill can express all these.

Conversely, our "Optician" algorithm can solve some classes of problems that
existing synthesis systems cannot.  In particular, techniques that only use
input-output examples are unable to synthesize programs with nested loops.  We
note that FlashFill's DSL does permit nested loops, but for efficiency reasons
the algorithm only synthesizes programs with no nesting (this is discussed at
the end of section 4.4 of "Automating String Processing in Spreadsheets Using
Input-Output Examples", POPL '11).

While a single nesting of loops seems to be sufficient for FlashFill's intended
uses, our benchmark suite (based on real-world uses of the Augeas tool) includes
many more complex formats and transformations.  In particular, these formats
tend to have deep layers of iterated data, requiring the synthesis of many
layers of nested loops.  FlashFill cannot handle such examples.  Our experiments
demonstrate that FlashFill only succeeds on 3 out of our 39 benchmarks, whereas
our system succeeds on all our benchmarks.

@@

     - What modifications did we need to make to examples in the Augeas and
       FlashFill benchmark suite to make them bijective?

We changed some formats in our benchmark suite to handle two forms of
non-bijectivity:
1) Whitespace present in one format but not the other.
2) Useful information projected away when going from one format to the other.

Category 1 was by far the most common, applying to many Augeas examples.  To
address this category, we added additional whitespace to the other format, where
typically whitespace is unnecessary.

@@

     - The restriction to bijective lenses is not fully justified

There are many formats kept in bijective correspondence with each other, like
those present at https://www.data.gov/ .

There are lens theories developed to handle synchronization tasks for data
formats not kept in bijective correspondence (quotient lenses to handle Category
1 non-bijectivity, and symmetric lenses to handle Category 2 non-bijectivity).
These theories have a bijective core, surrounded by either canonizers (for
quotient lenses) or projections (for symmetric lenses).  We believe that our
system can be used as a drop-in component for synthesis algorithms on these more
complex lens structures.

Reviewer 1 asked us directly about the potential for future work. Along those
lines, we have already implemented a system that synthesizes quotient lenses
from regular expressions augmented with equivalence relations to solve Category
1 non-bijectivity.  This system uses the algorithm provided in this paper as a
drop-in component without change. We conjecture the system could be augmented in
a similar way with projections as well.  We are not suggesting adding more
results of this form to the current paper because it is already very dense.

### 2017-09-14 papers/BijectiveLensSynth/popl_18_rebuttal.txt

     - What is the difference in expressiveness between our work and
        existing algorithms like FlashFill?

Formally, they are incomparable: each can synthesize some transformations
that the other cannot.

Since we restrict ourselves to bijections, we are unable to express functions
that don't preserve data, such as
  - mapping from a format with whitespace to one without whitespace,
  - turning a mix of uppercase and lowercase characters into purely
    uppercase characters, and
  - extracting a file name from a file path.
FlashFill can express all these.

Conversely, Optician can solve classes of problems that existing synthesis
systems cannot.  In particular, techniques that only use input-output examples
are currently unable to synthesize programs with nested loops. FlashFill's DSL
does permit nested loops, but for efficiency reasons the algorithm only
synthesizes programs with no nesting (see section 4.4 of "Automating String
Processing in Spreadsheets Using Input-Output Examples", POPL '11).

While a single nesting of loops is sufficient for FlashFill's intended uses, our
benchmark suite (based on real-world uses of the Augeas tool) includes many
complex formats and transformations.  In particular, these formats have many
layers of iterated data, requiring the synthesis of deeply nested loops.
FlashFill cannot handle such examples; FlashFill only succeeds on 3 of our
39 benchmarks, whereas our system succeeds on all 39 benchmarks.

@@

     - What modifications did we need to make to examples in the Augeas and
       FlashFill benchmark suite to make them bijective?

We modified some formats in our benchmark suite to handle two forms of
non-bijectivity:
1) Whitespace present in one format but not the other.
2) Useful information projected away when going from one format to the other.

Category 1 was by far the most common, applying to many Augeas examples.  To
address this category, we added whitespace to the other format where typically
whitespace is unnecessary.

Category 2 was less prevalent, but occurred, typically in the FlashFill
examples.  To address this category, we added the projected information to the
end of the format missing that information.

@@

     - The restriction to bijective lenses is not fully justified

There are many formats kept in bijective correspondence with each other, like
those present at https://www.data.gov/ .

Lens theories were developed to handle synchronization tasks for data
formats not kept in bijective correspondence (quotient lenses to handle Category
1 non-bijectivity, and symmetric lenses to handle Category 2 non-bijectivity).
These theories have a bijective core, surrounded by either canonizers (for
quotient lenses) or projections (for symmetric lenses).  We believe that our
system can be used as a drop-in component for synthesis algorithms on these more
complex lens structures.

Reviewer 1 asked us about the potential for future work. We have already
implemented a system that synthesizes quotient lenses from regular expressions
augmented with equivalence relations to solve Category 1 non-bijectivity.  This
system uses the algorithm provided in this paper as a drop-in component without
change. We conjecture Optician could be augmented in a similar way with
projections as well.  We are not suggesting adding more results of this form to
the current paper because it is already very dense.

     - After writing the regular expressions, is it very difficult for users to
       then write the lens?  Are we saving anybody significant effort?

@@

While we still require users to write regular expressions, programming
regular expressions is easier than programming lenses:

* Regular expressions are widely understood.  Few programmers know about
  programming with lenses.

* A regular expression describes one data format, while a lens deals with two
  at the same time.

### 2017-09-14 papers/BijectiveLensSynth/popl_18_rebuttal.txt

Thanks for your comments!

We begin with brief responses to the most significant concerns, followed by a
(long and optional) appendix addressing individual questions.

@@

     - What is the difference in expressiveness between our work and
        existing algorithms like FlashFill?

Formally, they are incomparable: each can synthesize some transformations
that the other cannot.

Since we restrict ourselves to bijections, we are unable to express functions
that don't preserve data, such as
  - mapping from a format with whitespace to one without whitespace,
  - turning a mix of uppercase and lowercase characters into purely
    uppercase characters, and
  - extracting a file name from a file path.
FlashFill can express all these.

Conversely, Optician can solve classes of problems that existing synthesis
systems cannot.  Techniques that only use input-output examples are currently
unable to synthesize programs with nested loops. FlashFill's DSL does permit
nested loops, but for efficiency reasons the algorithm only synthesizes programs
with no nesting ("Automating String Processing in Spreadsheets Using
Input-Output Examples", section 4.4, POPL '11).

While a single nesting of loops is sufficient for FlashFill's intended uses, our
benchmark suite (based on real-world uses of the Augeas tool) includes highly
complex formats and transformations.  In particular, these formats have many
layers of iterated data, requiring the synthesis of deeply nested loops.
FlashFill cannot handle such examples; FlashFill only succeeds on 3 of our
39 benchmarks, whereas our system succeeds on all 39 benchmarks.

@@

     - What modifications did we need to make to examples in the Augeas and
       FlashFill benchmark suite to make them bijective?

We modified some formats in our benchmark suite to handle two forms of
non-bijectivity:
1) Whitespace present in one format but not the other.
2) Useful information projected away when going from one format to the other.

Category 1 was by far the most common, applying to many Augeas examples.  To
address this category, we added whitespace to the other format where typically
whitespace is unnecessary.

Category 2 was less prevalent, but occurred, typically in the FlashFill
examples.  To address this category, we added projected information to the
end of the format missing that information.

@@

     - The restriction to bijective lenses is not fully justified

There are many formats kept in bijective correspondence with each other, like
those present at https://www.data.gov/ .

Lens theories were developed to handle synchronization tasks for data
formats not kept in bijective correspondence (quotient lenses to handle Category
1 non-bijectivity, and symmetric lenses to handle Category 2 non-bijectivity).
These theories have a bijective core, surrounded by either canonizers (for
quotient lenses) or projections (for symmetric lenses).  We believe that our
system can be used as a drop-in component for synthesis algorithms on these more
complex lens structures.  Indeed, we have implemented a system that synthesizes
quotient lenses from annotated regular expressions to solve Category 1
non-bijectivity,  using the algorithm provided in this paper as a drop-in
component without change.  We are not suggesting adding more results of this
form to the current paper because it is already very dense.

@@

     - After writing the regular expressions, is it very difficult for users to
       then write the lens?  Are we saving anybody significant effort?

While we require users to write regular expressions, programming regular
expressions is easier than programming lenses:
